---
title: "Untitled"
output: 
  pdf_document: 
    keep_tex: yes
---

```{r Prblm8Chp2}
# Problem 8 from Chapter 2

# a) Read "college" dataset into R ----
url <- 'http://www-bcf.usc.edu/~gareth/ISL/College.csv'
setInternet2()
college <- read.csv(url)

# b) ---- 
# College names are already row names. Will not do code to transform that
# will replace "fix()" with "head()" for printing
#fix(college)
head(college)
rownames(college) <- college[, 1]
#fix(college)
head(college)
college <- college[, -1]
#fix(college)
head(college)

# c) ----
# i.
summary(college)

# ii. scatterplot matrix of first 10 columns
pairs(college[, 1:10])

# iii. Create boxplot of outstate vs private
plot(college$Private, college$Outstate)
title('C) iii. Box Plot of Outstate vs Private',
      xlab = 'College Is Private',
      ylab = 'Number of Out of State Students')

# iv. Create qualitative variable Elite
elite <- rep('no', nrow(college))
elite[college$Top10perc > 50] = 'yes'
elite <- as.factor(elite)
college <- data.frame(college, elite)

summary(college$elite)
plot(college$elite, college$Outstate)
title('C) iv. Box Plot of Outstate vs Elite',
      xlab = 'College Is Elite',
      ylab = 'Number of Out of State Students')


# v. produce histograms with differing numbers of bins
vars <- names(college)
factVars <- vars[vapply(college, class, character(1)) == 'factor']

par(mfrow = c(3, 3))

hist(college$Apps, breaks = 35)
hist(college$Accept, breaks = 35)
hist(college$Enroll, breaks = 35)
hist(college$Room.Board, breaks = 40)
hist(college$Book, breaks = 50)
hist(college$Personal, breaks = 50)
hist(college$Grad.Rate, breaks = 50)
hist(college$Top10perc, breaks = 25)
hist(college$Top25perc, breaks = 25)

# vi. Continue to explore the data and write what you find.
```

From the previous question we can see various things from the histograms.
In the first row of the plots, the distribution for the number of applications,
accepted students, and student enrollment are all skewed right. This means that
only a few colleges get a large number of applications, accept a large number of students,
and enroll a large number of students. Further analysis will go into seeing if it is the same 
colleges are "outliers" compared to the rest of the data, which may be an indicator of a data
issue. In the second row, we can see various distributions of expenditures per college. In each
graph, there is skewness to the right with the largest skewness in the Book cost and Personal cost.
Interestingly, Room and Board costs are almost normally distributed, meaning that it may be independent
from book and personal costs and can be confirmed by the following scatter plots due to the lack of
correlation between the variables.

```{r}
pairs(college[, c('Room.Board', 'Books', 'Personal')])
```

Lastly, looking at the final row of the histograms, we can see several things. First there is a data
issue in the dataset since in the plot of Grad.Rate, one school has a rate greater than 100%,
which is impossible. Besides that the data are almost normal but have a little skewness left, 
meaning that although there is a little skewness, since there is a normal distribution of graduation
rate, we can make a hypothesis that there is not grade inflation. Moving on to the high school 
achievement, it looks like the new students from the to 10 percent is skewed right and only a few 
schools are predominantly made of up these students.  On the other hand, when new students are
from the top 25% of their class, the data is more normal. Overall, we can see that there are a few
elite schools that are made up of students that are only from the top 10 percent but otherwise, the
rest of the colleges in the dataset make up a normal distribution of the country based on high school
education since the graph seems to be normally distributed around 50%.
Percent of their class. For the graduation rate, 