---
title: "homework4"
author: "Jonathan Hamburg"
date: "November 8, 2016"
output: 
  pdf_document: 
    keep_tex: yes
---


```{r setup}
library(caret)
library(np)
library(kknn)
library(splines)
library(ISLR)
library(gridExtra)
library(parallel)
library(dplyr)
library(tidyr)
wage <- Wage

```


```{r probA}
x <- wage$age
y <- wage$wage
xyDF <- data.frame(x = x, y = y)

agelims <- range(x)
age_grid <- seq(from = agelims[1], to = agelims[2])
```

```{r qa1}

bandwidthByCV <- function(x, y, h = NULL, nfold = 5, 
                          type = c('kernel', 'knn', 'lpf', 'splines'),
                          seed = NULL, kerType = NULL, numBWs = 50) {
  # Load require packages
  require(caret)
  require(np)
  require(kknn)
  require(splines)
  
  if (missing(type)) {
    stop('Missing type parameter')
  } else {
    stopifnot(type %in% c('kernel', 'knn', 'lpf', 'splines'))
    if (length(type) > 1 ) {
      multiRun <- TRUE
    } else {
      multiRun <- FALSE
    }
  }
  
  # Argument logic
  stopifnot(length(x) > 1, length(x) == length(y), nfold == trunc(nfold),
            nfold <= length(x))

  if (!is.null(seed)) {
    stopifnot(is.numeric(seed))
    set.seed(seed)
  } 
  
  if (is.null(kerType)) {
    kerType <- 'gaussian'
  } else {
    kerTypes <- c('gaussian', 'epanechnikov', 'uniform')
    stopifnot(is.character(kerType) && kerType %in% kerTypes)
  }

    # If bandwiths aren't specified, add defaults
  if (is.null(h)) {
    h <- vector('list', length(type))
    for (i in 1:length(type)) {
      h[[i]] <- 
      switch(type[i],
             'kernel' = seq(from = 0.1, to = 5, length.out = numBWs),
             'knn' = {
               end <- floor(.10 * length(x))
               res <- seq(from = 3, to = end, length.out = numBWs)
               floor(res)
             },
             'lpf' = seq(from = 0.2, to = 1.5, length.out = numBWs),
             'splines' = seq(from = 0.01, to = 2, length.out = numBWs))    
    }
  } 

  # Create an empty matrix to fill in
  foldMSEs <- vector('list', length(type))
  for (i in 1:length(type)) {
    foldMSEs[[i]] <- matrix(NA, nrow = nfold, ncol = length(h[[i]]))
    colnames(foldMSEs[[i]]) <- h[[i]]
  }

  # Create folds for cv
  n <- length(x)
  flds <- caret::createFolds(x, k = nfold, list = TRUE, returnTrain = FALSE)

  # For each fold and bandwidth --- calculate MSE
  for (i in seq_along(flds)) {
    fold <- flds[[i]]
    trainX <- x[-fold]
    trainY <- y[-fold]
    train <- data.frame(x = trainX, y = trainY)
    testX <- x[fold]
    testY <- y[fold]
    test <- data.frame(x = testX, y = testY)

    if (i %% 20 == 0) {
      cat('On the ', i, 'iteration.\n')
    }

    for (j in 1:length(type)) {
      
      for (n in seq_along(h[[j]])) {
        bw <- h[[j]][n]
        
        foldMSEs[[j]][i, n] <- 
          switch(type[[j]],
                 'kernel' = { 
                   mod <- np::npreg(bws = bw, txdat = trainX, tydat = trainY,
                         exdat = testX, eydat = testY, bandwidth.compute = FALSE,
                         ckertype = kerType)
                   mod$MSE
                 },
                 'knn' = {
                   mod <- kknn::kknn(y ~ x, train, test, k = bw, 
                                     kernel = 'rectangular')
                   mse <- sum((mod$fitted.values - testY) ^ 2) / length(testY)
                   mse
                 },
                 'lpf' = {
                   # Need to update how ran if test data has points outside of
                   mod <- loess(y ~ x, train, span = bw, model = TRUE,
                                degree = 1, surface = 'direct')
                   
                   pred <- predict(mod, testX) 
                   mse <- sum((pred - testY) ^ 2) / length(testY)
                   mse
                 },
                 'splines' = {
                   mod <- smooth.spline(x = trainX, y = trainY, spar = bw)
                   pred <- predict(mod, testX)
                   mse <- sum((pred$y - testY) ^ 2) / length(testY)
                   mse
                 })
      }
    }
  }
  
  # Find average MSE for each bandwidth and then find the best one
  names(foldMSEs) <- type
  
  mseCV <- lapply(foldMSEs, colMeans)
  names(mseCV) <- type
  
  bestK <- lapply(mseCV, function(x) x[which.min(x)])
  names(bestK) <- type
  
  res <- list('bestBandwidth' = bestK,
              'cvMSEs' = mseCV,
              'foldMSEs' = foldMSEs)
  res
}
```

```{r 2a}

# Find the Best Parameter

kern <- bandwidthByCV(x, y, nfold = 5, type = 'kernel', seed = 10)
knn <- bandwidthByCV(x, y, nfold = 5, type = 'knn', seed = 10)
lpf <- bandwidthByCV(x, y, nfold = 5, type = 'lpf', seed = 10)
splin <- bandwidthByCV(x, y, nfold = 5, type = 'splines', seed = 10)


bestBW <- data.frame('kernel' = as.numeric(names(kern$bestBandwidth$kernel)),
                     'KNN' = as.numeric(names(knn$bestBandwidth$knn)),
                     'LocalPoly' = as.numeric(names(lpf$bestBandwidth$lpf)),
                     'Splines' = as.numeric(names(splin$bestBandwidth$splines)))

bestBW <- data.frame(lapply(bestBW, function(x) round(x, 4)))
bestBW
```

```{r 2b}
# 2 B

# Find the Best Parameter

kernMod <- np::npreg(bws = bestBW$kernel, txdat = x, tydat = y, 
                     bandwidth.compute = FALSE)
knnMod <- kknn::kknn(y ~ x, xyDF, data.frame(x = age_grid),
                     k = bestBW$KNN, kernel = 'rectangular')
lpfMod <- loess(y ~ x, xyDF, span = bestBW$LocalPoly, 
                model = TRUE, degree = 1, surface = 'direct')
splinMod <- smooth.spline(x = x, y = y, spar = bestBW$Splines)


# Get fitted values of age_grid
kernFit <- predict(kernMod, age_grid, se.fit = FALSE)
knnFit <- fitted(knnMod)
lpfFit <- predict(lpfMod, age_grid)
splinFit <- predict(splinMod, age_grid)$y


basePlot <- ggplot() + geom_point(aes(x, y)) 

kernPlotFit <- 
  basePlot + 
  geom_line(aes(age_grid, kernFit), col = 'red',
            lwd = 1.5) +
  ggtitle(paste('Kernel with bandwidth', bestBW$kernel))

knnPlotFit <- 
  basePlot + 
  geom_line(aes(age_grid, knnFit), col = 'red',
            lwd = 1.5) +
  ggtitle(paste('KNN with bandwidth', bestBW$KNN))


lpfPlotFit <- 
  basePlot + 
  geom_line(aes(age_grid, lpfFit), col = 'red',
            lwd = 1.5) +
  ggtitle(paste('Local Poly with bandwidth', bestBW$LocalPoly))


splinPlotFit <- 
  basePlot + 
  geom_line(aes(age_grid, splinFit), col = 'red',
            lwd = 1.5) +
  ggtitle(paste('Splines with bandwidth', bestBW$Splines))

grid.arrange(kernPlotFit, knnPlotFit, lpfPlotFit, splinPlotFit,
             newpage = FALSE)


```

```{r 2c}
# 2 C: Plot Residuals against fitted values


# Get fitted values of wage$age
kernFitY <- predict(kernMod, x, se.fit = FALSE) 

knnMod1 <- kknn::kknn(y ~ x, xyDF, data.frame(x = x),
                     k = bestBW$KNN, kernel = 'rectangular')

knnFitY <- fitted(knnMod1)
lpfFitY <- predict(lpfMod, x)
splinFitY <- predict(splinMod, x)$y

kernResid <- y - kernFitY
knnResid <- y - knnFitY
lpfResid <- y - lpfFitY
splinResid <- y - splinFitY

kernPlotFitResid <- 
  ggplot() + 
  geom_point(aes(kernFitY, kernResid)) +
  ggtitle(paste('Kernel with bandwidth', bestBW$kernel))

knnPlotFitResid <- 
  ggplot() + 
  geom_point(aes(knnFitY, knnResid)) +
  ggtitle(paste('KNN with bandwidth', bestBW$KNN))


lpfPlotFitResid <- 
  ggplot() + 
  geom_point(aes(lpfFitY, lpfResid)) +
  ggtitle(paste('Local Poly with bandwidth', bestBW$LocalPoly))


splinPlotFitResid <- 
  ggplot() + 
  geom_point(aes(splinFitY, splinResid)) +
  ggtitle(paste('Splines with bandwidth', bestBW$Splines))

grid.arrange(kernPlotFitResid, knnPlotFitResid, 
             lpfPlotFitResid, splinPlotFitResid,
             newpage = FALSE)

```


```{r q2 extraStuff, eval=FALSE, include=FALSE}

# 
# kernelFunc <- function(bw, trX, trY, tsX = NULL, tsY = NULL) {
#   
#   if ((is.null(tsX) & !is.null(tsY)) || (!is.null(tsX) & is.null(tsY))) {
#     stop('Either tsX and tsY must both be NULL or both have')
#   }
#   
#   stopifnot(is.numeric(trx), is.numeric(trY), length(trx) == length(trY),
#             length(bws) == 1, is.numeric(bws))
#   
#   if (!is.null(tsX)) {
#     stopifnot(length(tsX) == length(tsY))
#     
#     np::npreg(bws = bw, txdat = trX, tydat = trY, 
#               exdat = tsX, eydat = tsY)
#   } else {
#     np::npreg(bws = bw, txdat = trX, tydat = trY)
#   }
# 
# }

# plot(x, y)
# lines(age_grid, kernFit, col = 'red', lwd = 3)
# lines(age_grid, knnFit, col = 'blue', lwd = 3)
# lines(age_grid, lpfFit, col = 'green', lwd = 3)
# lines(age_grid, splinFit, col = 'pink', lwd = 3)


# 
# par(mfrow = c(2, 2))
# 
# # Kernel
# plot(x, y)
# lines(age_grid, kernFit, col = 'red', lwd = 3)
# title(paste('Fit via Kernel with bandwidth', bestBW$kernel))
# 
# # KNN
# plot(x, y)
# lines(age_grid, kernFit, col = 'red', lwd = 3)
# title(paste('Fit via KNN with bandwidth', bestBW$kernel))
#  
# # Local Polynomial
# 
# 
# # Splines



```





```{r q3 }

runSmoothSim <- function(n, x_grid){
  
  # Generate Data
  set.seed(10)
  x1 <- runif(n, 0, 1)
  eps <- rnorm(n, 0, .4)
  y1 <- sin(2 * pi * (1 - x1 ^ 2)) + (x1 * eps)
  actY <- sin(2 * pi * (1 - x1 ^ 2))
  df <- data.frame(x1, eps, y1, actY)
  
  res <- bandwidthByCV(x1, y1, nfold = n, seed = 10,
                       type = c('kernel', 'lpf', 'splines'),
                       kerType = 'epanechnikov', numBWs = 30)

  bestBW <- data.frame('kernel' = as.numeric(names(res$bestBandwidth$kernel)),
                       'LocalPoly' = as.numeric(names(res$bestBandwidth$lpf)),
                       'Splines' = as.numeric(names(res$bestBandwidth$splines)))
  
  bestBW <- data.frame(lapply(bestBW, function(x) round(x, 4)))
  bestBW
  
  kernMod <- np::npreg(bws = bestBW$kernel, txdat = x1, tydat = y1, 
                       bandwidth.compute = FALSE)
  lpfMod <- loess(y1 ~ x1, df, span = bestBW$LocalPoly, 
                  model = TRUE, degree = 1, surface = 'direct')
  splinMod <- smooth.spline(x = x1, y = y1, spar = bestBW$Splines)
  
  
  # Get fitted values of age_grid
  kernFitGrid <- predict(kernMod, x_grid, se.fit = FALSE)
  lpfFitGrid <- predict(lpfMod, x_grid)
  splinFitGrid <- predict(splinMod, x_grid)$y
  
  graphFits <- 
    data.frame('kernel' = kernFitGrid, 'lpf' = lpfFitGrid,
               'spline' = splinFitGrid, x = x_grid) %>%
    tidyr::gather(model, y, -x)
  
  # Plots
  basePlot <- ggplot() + geom_point(aes(x1, y1), alpha = .4) 
  
  finalPlot <- 
    basePlot + 
    geom_line(aes(x, y, color = model), data = graphFits, lwd = 1.5) +
    ggtitle(paste('N = ', n,
                  '\nKernel with bandwidth', bestBW$kernel, 
                  '\nLocal Poly with bandwidth', bestBW$LocalPoly,
                  '\nSplines with bandwidth', bestBW$Splines))
  
  # Distance Measures
  kernFit <- predict(kernMod, x1, se.fit = FALSE)
  lpfFit <- predict(lpfMod, x1)
  splinFit <- predict(splinMod, x1)$y
  
  kernMSE <- sum((kernFit - y1) ^ 2) / n
  lpfMSE <- sum((lpfFit - y1) ^ 2) / n
  splinMSE <- sum((splinFit - y1)) ^ 2 / n
  
  kernMSEAct <- sum((kernFit - actY) ^ 2) / n
  lpfMSEAct <- sum((lpfFit - actY) ^ 2) / n
  splinMSEAct <- sum((splinFit - actY) ^ 2) / n
  
  models <- list('kern' = kernMod, 'lpf' = lpfMod, 'splin' = splinMod)
  fits <- list('kern' = kernFit, 'lpf' = lpfFit, 'splin' = splinFit)
  mses <- list('kern' = kernMSE, 'lpf' = lpfMSE, 'splin' = splinMSE)
  msesAct <- list('kern' = kernMSEAct, 'lpf' = lpfMSEAct, 'splin' = splinMSEAct)
  
  return(list('bandwidths' = bestBW,
              'sPlot' = finalPlot,
              'plotData' = graphFits,
              'models' = models,
              'fits' = fits,
              'mses' = mses,
              'msesAct' = msesAct,
              'data' = df))
}

# X Data for predicting on the models
x_grid <- seq.int(from = 0, 1, length.out = 100)

# vector of N amounts --- will change it up parallel processing
amts <- seq(100, 1500, by = 100)
ramts <- rev(amts)
upd_amts <- c(amts, ramts)

for (i in 1:length(amts)) {
  upd_amts[i * 2 - 1] <- amts[i]
  upd_amts[i * 2] <- ramts[i]
}
upd_amts <- upd_amts[1:length(amts)]

```

```{r Set up parallel processing, include=FALSE}

# For parallel processing due to number of simulations
numCores <- detectCores()
cl <- parallel::makeCluster(numCores)

clusterEvalQ(cl, library(caret))
clusterEvalQ(cl, library(np))
clusterEvalQ(cl, library(kknn))
clusterEvalQ(cl, library(splines))
clusterEvalQ(cl, library(ggplot2))
clusterEvalQ(cl, library(dplyr))
clusterEvalQ(cl, library(tidyr))

clusterExport(cl, c('x_grid', 'runSmoothSim', 'bandwidthByCV', 'upd_amts'))

```

```{r Run simulation}
# Run the simulation

res <- vector('list', length(amts))
res <- parLapply(cl, upd_amts, function(n) runSmoothSim(n, x_grid))

stopCluster(cl)

```


```{r q3 convergence}

names(res) <- upd_amts

ord <- order(upd_amts)
resOrd <- res[ord]

msesList <- lapply(resOrd, function(x) as.data.frame(x$mses))
mses <- bind_rows(msesList)
mses$n <- as.numeric(names(resOrd))

# Distance based on the full function including error term
mses

msesActList <- lapply(resOrd, function(x) as.data.frame(x$msesAct))
msesAct <- bind_rows(msesActList)
msesAct$n <- as.numeric(names(resOrd))

# Distance based on the function without the error term
msesAct

```

Through this simulation, it looks like the Splines model converged the
fastest. We can conclude that the function converges on the 4th iteration
as the it does not reach above .002 after that. With this same logic, the 
Local Polynomial model reaches convergence at about the 7th iteration. The
Epanechnikov Kernel, technically never reaches convergence using that same
logic, however, it does rather stabilize at around the 11th iteration in that 
it doesn't change higher than .032 after that iteration.  

This is consistent with the graphs below, which show the different models
at the different number of data points. At around the 4th iteration, the splines
model pretty much looks similar in all subsequent graphs. The Local Polynomial 
settles down at about the 7th model, which we can tell by comparing it to the 
Splines model. As noted above, the Kernel model doesn't ever really converge to the
true function, however, it does stop moving by about the 11th iteration.

Below are the graphs of the models over the different iterations:

```{r q3 graphs}

# Graphs
lapply(resOrd, function(x) x$sPlot)


# plts <- lapply(resOrd, function(x) x$sPlot)
# names(plts) <- paste0('plot', as.numeric(names(resOrd)))
# 
# marrangeGrob(plts, ncol = 1, nrow = 2)
```

